---
title: "JSE Delistings and Stock Prices"
author: "Elias Muchineripi Mashayamombe"
date: "`r Sys.Date()`"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Introduction

This paper predicts the stock prices for 19 JSE listed companies and uses machine learning techniques to predict the prices of these stocks


```{r data, echo=FALSE, warning=FALSE, message=FALSE, results='hide'}
library(pacman)
pacman::p_load(data.table, fixest, BatchGetSymbols, finreportr, tidyverse, lubridate, ggplot2)

library(pacman)
pacman::p_load(data.table, fixest, BatchGetSymbols, finreportr, tidyverse, lubridate, ggplot2)

first.date <- Sys.Date() - 2500
last.date <- Sys.Date()
freq.data <- "monthly"
tickers <- c("WHL", "MTN", "SOL", "SHP", "BVT", 
             "IMP", "HAR", "DSY", "SNT", "AIP", "RMB",
             "STX", "ADR", "SBK", "CFR", "INL", "TFG", "PSG", "KBO")

# Get Stock Prices
stocks <- BatchGetSymbols(tickers = tickers, 
                          first.date = first.date,
                          last.date = last.date, 
                          freq.data = freq.data,
                          do.cache = FALSE,
                          thresh.bad.data = 0)

## Verify Returns
stocks_data <- stocks$df.tickers %>% setDT() %>%          # Convert to data.table
  .[order(ticker, ref.date)]   
head(stocks)
```

## Including Plots

You can also embed plots, for example:

```{r returns_plot, echo=FALSE, warning=FALSE, message=FALSE, results='hide'}
returns_plot_all <- ggplot(stocks_data[ticker %in% c("MTN", "SOL", "STX", "RMB")], 
                           aes(x= ref.date, y = ret.adjusted.prices, colour = ticker)) +
  geom_line() + theme_bw() + 
  labs(title = "", x = "Date", y= "Monthly Returns", subtitle = "") 
returns_plot_all

```


```{r prices_plot, echo=FALSE, warning=FALSE, message=FALSE, results='hide'}
price_plot_all <- ggplot(stocks_data[ticker %in% c("MTN", "SOL", "STX", "RMB")], 
                         aes(x= ref.date, y = price.close, colour = ticker)) +
  geom_line() + theme_bw() + labs(title = "", x = "Date", y= "Closing Price", subtitle = "")

price_plot_all
```
# Pre-processing the Data

Get the data for only the "price.close" column

```{r training, echo=FALSE, warning=FALSE, message=FALSE, results='hide'}
stocks_price_close <- stocks_data[, .(price.close), by = ticker]

library(caTools)

# Split the data into training and test data sets
split <- sample.split(stocks_price_close$price.close, SplitRatio = 0.7)
stocks_price_close_train <- stocks_price_close[split == TRUE, ]
stocks_price_close_test <- stocks_price_close[split == FALSE, ]

# Building the Model
# Fit a time series model using the auto.arima() function in the forecast library

library(forecast)

# Use the model to make predictions on the test data
stocks_price_close_train$price.close <- as.numeric(stocks_price_close_train$price.close)

auto.arima_model <- auto.arima(stocks_price_close_train$price.close,
                               seasonal = TRUE,
                               xreg = stocks_price_close_train$price.close)


auto.arima_pred <- forecast(auto.arima_model, xreg = stocks_price_close_test$price.close)
```

# Evaluating the Model

Calculate the mean absolute error (MAE)

```{r mae, echo=FALSE, warning=FALSE, message=FALSE, results='hide'}

auto.arima_pred$mean
stocks_price_close_test$price.close
mae <- mean(abs(auto.arima_pred$mean - stocks_price_close_test$price.close))
mae
```

# Concluding Remarks


# Appendix

1. "WHL": WHK Group Limited.
2. "MTN": MTN Group Limited.
3. "SOL": Sasol Limited.
4. "SHP": Shoprite Holdings Limited.
5. "BVT": Bavaria Industries Group Limited.
6. "IMP": Impala Platinum Holdings Limited.
7. "HAR": Harmony Gold Mining Company Limited.
8. "DSY": Distell Group Limited
9. "SNT": Santova Logistics Limited
10. "AIP": Afriplex (Pty) Ltd.
11. "RMB": Rand Merchant Bank (a division of FirstRand Bank Limited).
12. "STX": Satrix Limited.
13. "ADR": Adcock Ingram Holdings Limited.
14. "SBK": Standard Bank Group Limited.
15. "CFR": Clinix Health Group Limited.
16. "INL": Invicta Holdings Limited.
17. "TFG": The Foschini Group Limited.
18. "PSG": PSG Group Limited.
19. "KBO": Kalon Venture Capital Proprietary Limited.
